---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---


## Loading and preprocessing the data
```{r}
setwd("/Users/esteebie/Documents/DATA/Courses/4. JHU/5. Reproducible Research/Project 1 - fitbit/RepData_PeerAssessment1-master")
data <- read.csv("activity.csv")
```

## What is mean total number of steps taken per day?
Remove NAs amd then remove days with 0 steps - since these correspond to the removal of NAs.
```{r}
library(ggplot2)
library(reshape2)
totals <- (tapply(data$steps,data$date,sum, na.rm=TRUE))
totals <- totals[totals !=0]
dates <- names(totals)
totals_df <- cbind(as.data.frame(totals),dates)
ggplot(totals_df, aes(x=dates, y=totals)) +geom_bar(stat="identity")
ggplot(totals_df,aes(x=totals)) +geom_histogram(bins=60)
```
The first plot shows the number of steps taken each day in sequence.
The second plot shows the number of days that a certain number of steps were taken on. The most frequent number of steps attained on any given day is around 10,000.

```{r}
summary(totals_df$totals)[3:4]
```
These are the mean and median numbers of steps taken per day.

## What is the average daily activity pattern?

```{r}
averages <- tapply(data$steps,data$interval,mean,na.rm=TRUE)
averages <- averages[averages !=0]
interval <- names(averages)
averages_df <- cbind(as.data.frame(averages),interval)
averages_df$interval <- as.numeric(as.character(averages_df$interval))
ggplot(averages_df, aes(x=interval,y=averages, group=1)) +geom_line()
which.max(averages)
```
The 835th interval has the highest average (206 steps)

## Imputing missing values
```{r}
colSums(is.na(data))
```
2304 rows have missing values. Impute these by taking the average across all non-NA intervals.

```{r}
imputed <- data
avg_int <- mean(data$steps, na.rm=TRUE)
for (i in 1:dim(imputed)[1]) {
  if (is.na(imputed$steps[i])) {
    imputed$steps[i] <- avg_int
  }
}

colSums(is.na(imputed))
```


```{r}
imp_totals <- (tapply(imputed$steps,imputed$date,sum, na.rm=TRUE))
imp_totals <- imp_totals[imp_totals !=0]
dates <- names(imp_totals)
imp_totals_df <- cbind(as.data.frame(imp_totals),dates)
ggplot(imp_totals_df, aes(x=dates, y=imp_totals)) +geom_bar(stat="identity")
ggplot(imp_totals_df,aes(x=imp_totals)) +geom_histogram(bins=60)
```

```{r}
summary(imp_totals_df$imp_totals)[3:4]
```

Using this method of imputing, the mean remains the same but the median value has increased slightly from 10765 to 10766 (and is now the same as the mean).


## Are there differences in activity patterns between weekdays and weekends?
```{r}
imputed$date <- as.Date(imputed$date)
imputed$wkdy <- as.factor(ifelse(weekdays(imputed$date) %in% c("Saturday","Sunday"),"weekend","weekday"))

weekends <- tapply(imputed$steps[imputed$wkdy=="weekend"],imputed$interval[imputed$wkdy=="weekend"],mean)
weekdays <- tapply(imputed$steps[imputed$wkdy=="weekday"],imputed$interval[imputed$wkdy=="weekday"],mean)

interval <- names(weekdays)

weekend <- as.vector(weekends)
weekday <- as.vector(weekdays)

df <- as.data.frame(cbind(weekend, weekday))
df <- cbind(interval, df)

df_mlt <- melt(df,id="interval", measure.vars=c("weekend","weekday"))
colnames(df_mlt) <- c("interval","day_type","steps")

df_mlt$interval <- as.numeric(as.character(df_mlt$interval))

ggplot(df_mlt, aes(x=interval, y=steps, facet=day_type, group=1, colour=day_type)) +geom_line() +facet_grid(day_type~.) +scale_x_continuous(breaks=seq(0,2400,200))
```

